---
output: github_document
fig_caption: yes
always_allow_html: true
---


```{r setup, out.width='100%', include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, 
                      comment = "#>", 
                      message = FALSE, 
                      warning = FALSE, 
                      eval = TRUE,
                      echo = FALSE)
library("tidyverse")
library("lubridate")
library("readxl")
library("tibble")
library("dataDownloader")
library("kableExtra")


theme_set(theme_bw(base_size = 12))


# Data
source("R/data_dic/data_dic.R")

biomass <- read_csv("data/biomass/FunCaB_clean_biomass_2015-2021.csv")
community <- read_csv("data/community/FunCaB_clean_composition_2015-2018.csv")
#cflux <- read_csv(file = "data/cflux/FunCaB_clean_Cflux_2015-2017.csv")


meta <- biomass %>% 
  distinct(siteID) %>% 
  mutate(biogeographic_zone = recode(siteID,
                             Ulvehaugen = "alpine",
                             Lavisdalen = "alpine",
                             Gudmedalen = "alpine",
                             Skjelingahaugen = "alpine",
                             Alrust = "sub.alpine",
                             Hogsete = "sub.alpine",
                             Rambera = "sub.alpine",
                             Veskre = "sub.alpine",
                             Fauske = "boreal",
                             Vikesland = "boreal",
                             Arhelleren = "boreal",
                             Ovstedalen = "boreal"),
        prep_level = recode(siteID,
                               Ulvehaugen = 1,
                               Lavisdalen = 2,
                               Gudmedalen = 3,
                               Skjelingahaugen = 4,
                               Alrust = 1,
                               Hogsete = 2,
                               Rambera = 3,
                               Veskre = 4,
                               Fauske = 1,
                               Vikesland = 2,
                               Arhelleren = 3,
                               Ovstedalen = 4),
        climate = paste0(biogeographic_zone, prep_level),
        climate = factor(climate, levels = c("boreal1", "boreal2", "boreal3", "boreal4", "sub.alpine1", "sub.alpine2", "sub.alpine3", "sub.alpine4", "alpine1", "alpine2", "alpine3", "alpine4")))

```

This is the git repository for the FunCaB project and base for the data paper:
Vandvik et al. (2022). Data descriptor: The role of functional group interactions in mediating climate change impacts on the Carbon dynamics and biodiversity of alpine ecosystems.



## Summary

Removal experiments allow disentangling of direct effects of climate from indirect effects, resulting from changes in interactions among species or functional groups with climate.
Across the Vestland Climate Grid (VCG), twelve sites in a climate grid spanning 4°C in growing season temperature and 2000 mm in mean annual precipitation in Western Norway, we conducted a fully factorial plant functional group removal experiment (graminoids, forbs, bryophytes).
Over six years, we recorded biomass, soil microclimate, plant community composition, seedling recruitment, ecosystem carbon fluxes, and reflectance in 384 experimental treatment and control plots.
The datasets consists of 5215 biomass records, 1,084,970 soil temperature records, 4,412 soil moisture records, 14,044 plant records covering 191 taxa,  3,696  carbon fluxes measurements, and XXX reflectance measurements.
The data can be combined with other data collected within the VCG on longer-term climate data and plant population, community, ecosystem, and functional trait data. 


## Datasets

```{r datasets}

tibble(Number = c("i", "ii", "", "iii", "iv", "v", "vi"),
       Variable = c("Biomass removal", "Soil microclimate", "", "Plant community composition", "Seedling recruitment", "Ecosystem carbon fluxes", "Reflectance"),
       Period = c("2015 - 2021", "2015 - 2018", "2015 - 2016", "2015 - 2019", "2018 - 2019", "2015 - 2017", "2019 - 2021"),
       Filename = c("FunCaB_clean_biomass_2015-2021.csv", "FunCaB_clean_soilMoisture_2015-2018.csv", "FunCaB_clean_soiltemperature_2015-2016.csv", "FunCaB_clean_composition_2015-2018.csv", "FunCaB...", "FunCaB_clean_Cflux_2015-2017.csv", "FunCaB...")) %>% 
  kbl()
```



**Project description**, **data dictionaries**, and the overview of all **data sets** are in the data paper (REF).

The **raw** and **clean data** from the project are stored and available on **OSF**: https://osf.io/4c5v2/

All **R code** to clean and curate the data is on **GitHub repository**: https://github.com/Between-the-Fjords/funcab_data



### Naming conventions across datasets

This is the naming convention used across the datasets and can be used to link the datasets.
These variables are also used in the datasets from the Vestland Climate Grid (VCG) and can be used to join these datasets.
Note that the turfID contains information about the control plots from the VCG.

| Data or file  | Description of naming convention  | Example  |
|:---|:---|:---|
| File names  | Project_status_Dataset_Year.extension  | FunCaB_clean_biomass_2015-2021.csv  |
| year  | Year of observation | 2015 |
| date  | Date of observation | yyyy-mm--dd |
| date_time  | Date and time of observation | yyyy-mm-dd hh-mm-ss |
| siteID  | Unique siteID | Vikesland |
| blockID  | Unique blockID as a combination of siteID and block number | Arh3  |
| plotID  | Unique plotID as a combination of siteID, block and treatment | Arh3FGB  |
| treatment  | Treatment of functional group removal indicating which groups have been removed; F = forbs, G = graminoids, B = Bryophytes | C, B, F, G, FB, GB, GF, FGB |
| turfID  | Unique turfID corresponding to the Vestland Climate Grid control plots | Arh3FGB  |
| species  | Vascular plant taxon names follow for Norway (Lid J & Lid, 2010). The full taxa is written using genus and species with a blank. | Rancunculus acris|
| responses  | response variable | cover, biomass, Reco, etc. |



## Data dictionaries

### Biomass removal

### Biomass


```{r biomass}

biomass_plot <-  biomass %>% 
  left_join(meta, by = c("siteID")) %>% 
   mutate(treatment = factor(treatment, levels = c("B", "F", "G", "FB", "GB", "GF", "FGB"))) %>% 
  ggplot(aes(x = treatment, y = biomass, fill = removed_fg)) +
  geom_col() +
  scale_fill_manual(name = "Functional group", values = c("orange", "purple", "limegreen")) +
  labs(x = "Removed functional group", y = "Biomass in g") +
  facet_grid(year ~ climate) +
  theme_bw() +
  theme(legend.position="top")

biomass_plot

```


```{r biomass-dic}
 knitr::kable(biomass_dic)
```


### Soil microclimate

```{r soiltemp-dic}
 knitr::kable(soil_temperature_dic)
```

```{r soilmoist-dic}
 knitr::kable(soilmoisture_dic)
```


### Plant community composition

``` {r community}

comm_data <- community %>%
  select(year, siteID:mossCov, -turfID) %>%
  filter(treatment %in% c("C", "FB", "GB", "GF"),
         year != 2016) %>%
  #filter(blockID == "Alr5") %>%
  distinct() %>%
  mutate(graminoidCov = if_else(year == 2015 & treatment %in% c("GB", "GF"), NA_real_, graminoidCov),
         forbCov = if_else(year == 2015 & treatment %in% c("FB", "GF"), NA_real_, forbCov),
         mossCov = if_else(year == 2015 & treatment %in% c("FB", "GB"), NA_real_, mossCov)) %>%
  group_by(year, siteID, treatment) %>%
  summarise_at(c("graminoidCov", "forbCov", "mossCov"), mean, na.rm = TRUE) %>%
  pivot_longer(cols = c(graminoidCov, forbCov, mossCov), names_to = "fg", values_to = "fg_cover") %>%
  pivot_wider(names_from = treatment, values_from = fg_cover) %>%
  mutate(Forbs = GB - C,
         Graminoids = FB - C,
         Bryophytes = GF - C) %>%
  select(-C, -FB, -GB, -GF) %>%
  pivot_longer(cols = c(Forbs, Graminoids, Bryophytes), names_to = "FunctionalGroup", values_to = "Cover") %>%
  filter(!is.na(Cover)) %>%
  mutate(FunctionalGroup = factor(FunctionalGroup, levels = c("Graminoids", "Forbs", "Bryophytes")))

community_plot <- comm_data %>%
  left_join(meta, by = "siteID") %>% 
  mutate(biogeographic_zone = factor(biogeographic_zone, levels = c("alpine", "sub.alpine", "boreal"))) %>% 
  ggplot(aes(x = year, y = Cover, colour = FunctionalGroup)) +
  geom_line() +
  geom_hline(yintercept = 0, colour = "grey") +
  scale_colour_manual(name = "Functional group", values = c("limegreen", "purple", "orange")) +
  facet_grid(biogeographic_zone ~ prep_level) +
  theme_set(theme_bw(base_size = 18))

community_plot



```

```{r community-dic}
 knitr::kable(community_dic)
```


### Ecosystem carbon fluxes

``` {r cflux}


cflux %>%
  filter(year == 2017) %>%
  left_join(meta, by = c("siteID")) %>%
  mutate(treatment = factor(treatment, levels = c("C", "B", "F", "G", "FB", "GB", "GF", "FGB")),
         biogeographic_zone = factor(biogeographic_zone, levels = c("alpine", "sub.alpine", "boreal"))) %>%
  ggplot(aes(y = gpp, treatment)) +
  geom_boxplot() +
  facet_grid(biogeographic_zone ~ prep_level)



```


```{r cflux-dic}
 knitr::kable(cflux_dic)
```
